---
title: "The Bestseller Palette: Data-Backed Trends in Book Cover Colors"
author: Andrea Rivera & Maria Carda 
date: today
format:
  html:
    theme: [style.scss]
    embed-resources: true
    toc: true
    toc-location: right
    toc-title: Index
editor: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=FALSE, warning=FALSE, knitr.purl.inline = TRUE )
```

#Introduction

Book cover design plays a crucial role in attracting readers and
influencing purchasing decisions. One key aspect of cover design is
color, which conveys mood, genre conventions, and marketing strategies.
This study aims to analyze the color palettes of book covers across
different genres in two major online book platforms: the top-selling
sections of *Casa del Libro* and compare them to *Goodreads Readers'
Favorite Books 2024*.

By examining whether specific color schemes are more prevalent in
certain categories, this research seeks to uncover patterns in book
cover aesthetics and their possible correlation with genre expectations
and consumer preferences.

To conduct this analysis, we used R for web scraping to collect cover
images from both platforms and extract dominant color palettes using
image processing techniques.

# Libraries

```{r}
#| message: false

rm(list = ls()) 

packages = c("RSelenium", "rvest", "magrittr", "tidyr", "dplyr", "magick", "tidyverse", "ggplot2", 
             "scales", "shiny", "writexl", "shinyjs", "shinythemes","shinyWidgets", "rsconnect")

package.check <- lapply(packages,
                        FUN = function(x){
                          if (!require(x,character.only = TRUE)){
                            install.packages(x,dependencies = TRUE)
                            library(x, character.only = TRUE)
                          }
                        }
)


```

# Casa del Libro

*Casa del Libro* is a well-established Spanish bookstore chain and
online platform that serves as a hub for book lovers. It provides a vast
catalog of books across various genres, offering both physical and
digital formats.

## Categories & sub-categories

For this research, we analyzed the best-selling books on *Casa del
Libro* across five key genres:

-   Biographies

-   Science Fiction

-   Mystery & Thriller

-   Romance

-   Young Adult Fantasy

To achieve this, we first identified the paths to the broader categories
that contain these subcategories. This step allows us to systematically
navigate the website and efficiently scrape the relevant data.

```{r}
# Categories & Subcategories
categories <- data.frame(
  category = c("Biography", "Science Fiction", "Mystery & Thriller", "Romance", 
               "Young Adult Fantasy"),
  category_xpath = c("//a[contains(text(),'No Ficción')]",
                     "//a[contains(text(),'Ficción')]", 
                     "//a[contains(text(),'Ficción')]", 
                     "//a[contains(text(),'Ficción')]", 
                     "//a[contains(text(),'Juvenil')]"),
  subcategory_xpath = c("//a[contains(text(),'Biografías')]",
                        "//a[span[contains(text(),'Novela de ciencia ficción')]]", 
                        "//a[span[contains(text(),'Novela negra')]]", 
                        "//a[span[contains(text(),'Novela romántica y erótica')]]",
                        "//a[span[contains(text(),'Fantasía')]]")
)
```

## Selenium

We will use Selenium to automate the scraping process, enabling us to
navigate *Casa del Libro’s* website and efficiently extract book data.
To achieve this, we initialize the Selenium remote driver and navigate
to the homepage, ensuring a seamless start to the data collection
process.

```{r}

remDr <- remoteDriver(port = 4449)
remDr$open()

#Go to Casa del Libro webpage
remDr$navigate("https://www.casadellibro.com")

Sys.sleep(5)

```

## Privacy window

If the privacy window appears, the code utilizes `tryCatch()` to manage
the rejection of the cookie consent banner on the *Casa del Libro*
website. It attempts to locate the "Reject All" button using an XPath
query. If the button is found, the code clicks on it to reject the
cookies. If the button is not present, the script continues without
interruption and outputs a message indicating that the privacy window
was not displayed. This approach ensures that the scraping process can
proceed seamlessly, regardless of whether the cookie consent banner is
shown.

```{r}

tryCatch({
  #Try to find the button
  rechazar <- remDr$findElement(using = "xpath", "//button[@id='onetrust-reject-all-handler']")
  
  # Click if finds it
  rechazar$clickElement()
  message("Cookie banner found and rejected.")
}, error = function(e) {
  
  # Do nothing if it does not find it
  message("No cookie banner detected. Continuing execution.")
})

```

## Scrapbook

### Function

The `scrap_books` function is designed to scrape book information from
the *Casa del Libro* website. It first navigates to the pre-defined
category and subcategory using XPath selectors. Once on the page, it
collects details such as book titles, authors, and image URLs from the
selected number of pages (in our case, 10 pages) using a combination of
CSS selectors and HTML parsing.

For each page, the function extracts the relevant book information,
storing it in separate lists for titles, authors, and image URLs. After
gathering the data, it attempts to navigate to the next page, clicking
the pagination button. If the button is not found, it stops the process.

Finally, the function compiles the extracted data into a dataframe that
includes the category name, book titles, authors, and image URLs, which
is then returned for further analysis or processing.

```{r}
# Scrapbook function
scrap_books <- function(category_xpath, subcategory_xpath, category_name, num_pages = 10) {
  
  # Click in category
  enlace <- remDr$findElement(using = "xpath", category_xpath)
  enlace$clickElement()
  
  Sys.sleep(5)
  
  # Click in subcategory
  enlace <- remDr$findElement(using = "xpath", subcategory_xpath)
  enlace$clickElement()
  
  # Create empty lists to store data
  all_titles <- c()
  all_authors <- c()
  all_img_urls <- c()

  # Extract data in n pages
  for (i in 1:num_pages) {
    
    Sys.sleep(5)  
    
    # Extract element of the book
    product_elems <- remDr$findElements(using = "css selector", "div.compact-product")
    product_html <- sapply(product_elems, function(x) x$getElementAttribute("outerHTML")[[1]])
    parsed_products <- lapply(product_html, read_html)
    
    # Titles
    titles <- sapply(parsed_products, function(x) x |> 
                       html_node("a.product-title") |> 
                       html_text())
    
    # Authors
    authors <- sapply(parsed_products, function(x) {
      author_node <- x |> 
        html_node("p.truncate-text")
      if (!is.null(author_node)) html_text(author_node) else NA
    })
    
    # Images' url
    img_urls <- sapply(parsed_products, function(x) {
      x %>% html_node("img") |> 
        html_attr("src")
    })
    
    # Save in empty lists
    all_titles <- c(all_titles, titles)
    all_authors <- c(all_authors, authors)
    all_img_urls <- c(all_img_urls, img_urls)
    
    # Go to next page
    boton_pagina <- tryCatch({
      remDr$findElement(using = "xpath", paste0("//button[normalize-space(text())='", i + 1, "']"))
    }, error = function(e) NULL)
    
    if (!is.null(boton_pagina)) {
      boton_pagina$clickElement()
      Sys.sleep(5)  
      remDr$refresh() 
    } else {
      message("Button ", i + 1, "not found.")
      break
    }
  }
  
  # Create dataframe with the data
  result <- data.frame(
    category = category_name,
    title = all_titles,
    author = all_authors,
    img_url = all_img_urls,
    stringsAsFactors = FALSE
  )
  
  return(result)
}

  
```

### Application

Once the function is defined, we apply it to scrape the data.

```{r}

# Apply function in each category & subcategory
temp_casa.libro <- list()

for (i in 1:nrow(categories)) {
  books <- scrap_books(
    category_xpath = categories$category_xpath[i],
    subcategory_xpath = categories$subcategory_xpath[i],
    category_name = categories$category[i]
  )
  
  temp_casa.libro[[categories$category[i]]] <- books
}
 

```

### Save results

Results are then save in a dataframe for further analysis.

```{r}
#save results
casa.libro <- do.call(rbind, temp_casa.libro)

```

## End selenium

Finally, we close Selenium after the scrapping process is complete. This
ensures that the browser window opened by Selenium is properly closed,
freeing up system resources and maintaining a clean environment for
subsequent tasks.

```{r}
#end selenium
remDr$close()

```

# Goodreads

*Goodreads* is an online platform that connects readers worldwide. It
offers features such as book reviews, ratings, and personalized
recommendations, fostering a vibrant community for book enthusiasts.
This includes the *Goodreads Choice Awards*, annual recognitions of the
year's most popular books across various genres that are chosen by the
platform's users.

For this research, we scraped the results of the 2024 awards, which
established a record-breaking number of user votes (6.2 million). The
selected categories to match as close as possible *Casa del Libro* were
the following:

-   Mystery and Thriller books
-   Romance books
-   Science Fiction books
-   History and Biography books
-   Young adult fantasy books

## Scrapping

The first step is to define the URLs that need to be scrapped and create
an empty dataframe where the data will be stored.

Then, a loop is created to go through each URL, retrieving the book
titles, cover image URLs, and category names. The extracted category
name is cleaned by removing the prefix "Readers' Favorite" and book
details are stored in a temporary data frame, which is then appended to
the main *Goodreads* data frame.

```{r}

url <- "https://www.goodreads.com/choiceawards/readers-favorite-mystery-thriller-books-2024"
url2 <- "https://www.goodreads.com/choiceawards/readers-favorite-romance-books-2024"
url3 <- "https://www.goodreads.com/choiceawards/readers-favorite-ya-fantasy-books-2024"
url4 <- "https://www.goodreads.com/choiceawards/readers-favorite-science-fiction-books-2024"
url5 <- "https://www.goodreads.com/choiceawards/readers-favorite-history-bio-books-2024"
urls <- c(url, url2, url3, url4, url5)

goodreads <- data.frame()

## Loop to extract the data
for (url in urls) {
    page <- read_html(url)
    
    # Extract book titles
    titles <- page %>%
        html_nodes(".pollAnswer__bookLink img") %>% 
        html_attr("alt")
    
    # Extract cover images
    covers <- page %>%
        html_nodes(".pollAnswer__bookLink img") %>% 
        html_attr("src")
    
    # Extract category from the HTML
    category <- page %>%
        html_element(".gcaMastheader") %>%  
        html_text() %>%  
        gsub("Readers' Favorite ", "", .) 
    
    # Create a temporary data frame
    temp_books <- data.frame(title = titles, img_url = covers, category = category, stringsAsFactors = FALSE)
    
    # Append to the main data frame
    goodreads <- bind_rows(goodreads, temp_books)
}


```

After collecting the book information, we extracted the book title and
author separately using regular expressions.

```{r}
# Use general expression to separate title and author
goodreads <- goodreads |> 
  rowwise() |> 
  mutate(author = str_extract(title, "(?<= by ).*"),
        title = str_extract(title, ".*(?= by )"))

```

# Join data

Once both webpages are harvested, we combined both data into a single
dataset and added a *webpage* variable to identify from which source is
each book from.

In addition, we recoded the category "History & Biography" from
*Goodreads* to simply "Biography" for consistency.

```{r}

best_books <- bind_rows(
  goodreads |> 
    mutate(webpage = "Goodreads", category = recode(category, "History & Biography" = "Biography")),
  casa.libro |> 
    mutate(webpage = "Casa del libro")
  )   

write_xlsx(best_books, "best_books.xlsx")

```

# Extract colors

## Function

Once we have our final dataset, we defined a new function
*extract_colors* to obtain color palettes from the book cover images
given its URL. It first reads the image from the provided URL and
resizes to reduce processing time. The image is then converted into
pixel data. If the image data is in bitmap format (type of image
representation where the image is made up of a grid of individual
pixels, each of which has its own color value), it is normalized to a
numeric range from 0 to 1 by dividing the pixel values by 255. The pixel
data is then converted into a data frame containing the RGB channels
(red, green, blue) for each pixel.

then, the function applies k-means clustering to the RGB data to
identify the most dominant colors in the image. Finally, it returns the
RGB values of the dominant colors identified by the clustering process
in a vector format that represents the colors.

```{r}

# Defined function
extract_colors <- function(image_url, n_colors = 6) {
  
  # read image from url
  img <- image_read(image_url)  |> 
    image_resize("50x50") # Resize to reduce processing time
  
  # Convert the image to pixel data
  img_data <- image_data(img)
  
  # Check if the structure is a bitmap
  if (inherits(img_data, "bitmap")) {
    # If it is a bitmap, we convert it to a numeric format
    img_data <- as.integer(img_data) / 255
  }
  
  # Convert the image to a data frame
  # Extract the RGB channels
  df <- data.frame(
    red = as.vector(img_data[,,1]),    # Red
    green = as.vector(img_data[,,2]),  # Green
    blue = as.vector(img_data[,,3])    # Blue
  )
  
  # Apply k-means to find dominant colors
  kmeans_result <- kmeans(df, centers = n_colors, nstart = 25, iter.max = 100)
  
  # Convert dominant colors to RGB format
  dominant_colors <- rgb(kmeans_result$centers[,1],
                         kmeans_result$centers[,2],
                         kmeans_result$centers[,3], maxColorValue = 1)
  
  # Return colors as a vector of colors 
  return(dominant_colors)
}

```

## Application

The function is then applied to our dataset

```{r}

best_books$colors <- sapply(best_books$img_url, extract_colors, simplify = FALSE)

```

## Visualization

Finally, we visualized the extracted colors by displaying a palette
derived from the first book's cover to ensure the previous steps worked
correctly.

```{r}

show_col(best_books$colors[[1]])

```

# Analysis

## Tidy data

To begin our analysis, each book's color palette (stored as a list) is
broken down into individual rows, making each color its own observation
allowing for more detailed color analysis.

```{r}
# Expand the dataset to individual rows for each color
best_books_expanded <- best_books |> 
  unnest(colors)

write_xlsx(best_books_expanded, "best_books_expanded.xlsx")
```

## Hex to RGB

Once the colors are unnested, we converted hex color codes into RGB
values to extract red, green, and blue components.

```{r}
# Function to convert hex colors to RGB
hex_to_rgb <- function(hex) {
  rgb_matrix <- col2rgb(hex) / 255  
  data.frame(hex = hex, R = rgb_matrix[1, ], G = rgb_matrix[2, ], B = rgb_matrix[3, ])
}

# Convert colors from hex to RGB
color_data <- best_books_expanded |> 
  mutate(rgb_values = map(colors, hex_to_rgb)) |> 
  unnest(rgb_values) 
```

## Clusters

Then, we applied k-means clustering to group colors into six clusters
per book category. If a category has fewer than six colors, unique
clusters are assigned to each row; otherwise, k-means clustering
organizes colors based on their RGB values. This approach helps identify
dominant color trends within different book genres.

```{r}
# k-means clustering to group colors into 6 clusters
set.seed(123) 

color_data_clustered <- color_data |> 
  group_by(category) |> 
  group_modify(~ {
    if (nrow(.x) < 6) {
      .x$cluster <- as.factor(seq_len(nrow(.x)))  
    } else {
      kmeans_result <- kmeans(.x[, c("R", "G", "B")], centers = 6)
      .x$cluster <- as.factor(kmeans_result$cluster)
    }
    return(.x)
  }) |> 
  ungroup()

```

### Cluster representative

To refine the previous clustering step, we chose to identify
representative color for each cluster. This provides a simplified way to
analyze dominant colors within each cluster.

```{r}
# Map clusters to representative colors 
cluster_representative <- color_data_clustered |> 
  group_by(category, cluster) |> 
  summarise(representative_color = first(hex))  

# Pick first as representative
```

### Map clusters

Following the cluster simplification, we joined the clusters and
representative colors to the expanded dataset.

```{r}
# Replace original colors with their cluster representative
best_books_clustered <- best_books_expanded |> 
    full_join(color_data_clustered, by = c("colors", "category", "title", "author", "webpage")) |> 
    full_join(cluster_representative, by = c("category", "cluster"))

```

## Representation of colors per category

Afterwards, we calculate the frequency of each color within each
category, rank them in descending order, and select the six most common
colors per category.

```{r}
# Count occurrences of grouped colors per category
color_counts <- best_books_clustered |> 
  group_by(category, representative_color) |> 
  summarise(Freq = n(), .groups = "drop") |> 
  arrange(category, desc(Freq))

top_colors_all <- color_counts |> 
  group_by(category) |> 
  arrange(desc(Freq)) |> 
  slice_head(n = 6) |> 
  mutate(position = row_number()) |> 
  ungroup()

```

### Color representation per category

```{r}
# Ensure text color contrasts with background
top_colors_all <- top_colors_all |> 
  mutate(text_color = ifelse(
    (col2rgb(representative_color)[1,] * 0.299 +
     col2rgb(representative_color)[2,] * 0.587 +
     col2rgb(representative_color)[3,] * 0.114) > 150, 
    "black", "white"
  ))


ggplot(top_colors_all, aes(x = factor(position), y = category, fill = representative_color)) +
    geom_tile(color = "gray70", linewidth = 0.7) +  
    scale_fill_identity() +  
    scale_color_identity() +  
    theme_minimal(base_size = 14) +
    labs(title = "Top 6 Most Frequent Colors by Category",
         subtitle = "Most dominant colors used in book covers across different genres",
         x = NULL, y = NULL) +
   theme(
    panel.background = element_rect(fill = "transparent", color = NA),  
    plot.background = element_rect(fill = "transparent", color = NA),   
    panel.grid = element_blank(), 
    axis.text.x = element_blank(),  
    axis.ticks = element_blank(),
    strip.text = element_text(face = "bold", size = 10), 
    axis.text.y = element_text(face = "bold", size = 10),  
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5)
)


```

### Color representation per category and webpage

Furthermore, we repeated the previous analysis for each webpage,
representing the color palette of each category per source. The
comparative analysis reveals notable differences in color usage between
the two platforms, indicating potential variations in design
preferences, marketing strategies, or audience expectations.

```{r}
cluster_representative <- color_data_clustered |> 
  group_by(category, cluster, webpage) |> 
  summarise(representative_color = first(hex)) 

best_books_clustered <- best_books_expanded |> 
    full_join(color_data_clustered, by = c("colors", "category", "title", "author", "webpage")) |> 
    full_join(cluster_representative, by = c("category", "cluster", "webpage"))

top_colors_all <- best_books_clustered |> 
  group_by(category, webpage, representative_color) |> 
  summarise(Freq = n(), .groups = "drop") |> 
  arrange(category, webpage, desc(Freq)) |>   # Ensure webpage is considered
  group_by(category, webpage) |> 
  slice_head(n = 6) |> 
  mutate(position = row_number()) |>  
  ungroup() |> 
  mutate(text_color = ifelse(
    grepl("^#", representative_color),  # Ensure it's a hex color
    ifelse(
      (col2rgb(representative_color)[1,] * 0.299 +
       col2rgb(representative_color)[2,] * 0.587 +
       col2rgb(representative_color)[3,] * 0.114) > 150, 
      "black", "white"
    ), 
    "black"  # Default to black if invalid color
  ))


ggplot(top_colors_all, aes(x = position, y = category, fill = representative_color)) +
    geom_tile(color = "white", size = 0.5) + 
    scale_fill_identity() +  
    scale_color_identity() +  
    theme_minimal() +
    labs(title = "Top 6 Most Frequent Colors by Category and Source", x = NULL, y = NULL) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        axis.text.y = element_text(face = "bold", size = 10)
    ) +
    facet_wrap(~webpage)  # Should separate by webpage



```

For example, *Science Fiction* on Casa del Libro features vibrant
purples, blues, and neon greens, whereas in Goodreads the palette
reflects subdued tones like gray and dark green. And in *Romance*, Casa
del Libro highlights deep reds and warm neutrals, while Goodreads
includes brighter blues and greens.

## Top author

In addition, we identified the most frequently occurring colors in the
book covers of the most prolific author in the dataset and visualized
the color palette. The script first determines the top author based on
the number of books, then extracts the six most dominant colors from
their book covers and each color is displayed as a tile in a bar-like
palette, with contrasting text labels for readability.

```{r}

top_author <- best_books %>% 
    count(author, sort = TRUE) %>% 
    arrange(desc(n)) %>%
    slice(1) %>%
    pull(author) %>% 
    .[[1]]  # Force extraction of only the first element


top_authors_colors <- best_books %>% 
    filter(author == top_author) %>% 
    unnest_longer(colors) %>% 
    count(author, colors, sort = TRUE)  %>% 
    slice_head(n = 6) %>% 
    mutate(rank = row_number())  %>% 
    ungroup()  %>%
  mutate(text_color = ifelse(
    grepl("^#", colors),  # Ensure it's a hex color
    ifelse(
      (col2rgb(colors)[1,] * 0.299 +
       col2rgb(colors)[2,] * 0.587 +
       col2rgb(colors)[3,] * 0.114) > 150, 
      "black", "white"
    ), 
    "black"  # Default to black if invalid color
  ))

ggplot(top_authors_colors, aes(x = rank, y = reorder(author, -n), fill = colors)) +
    geom_tile(color = "white", size = 0.5) + 
    geom_text(aes(label = colors, color = text_color), size = 3, fontface = "bold") +  
    scale_fill_identity() +  
    scale_color_identity() +   
    theme_minimal() +
    labs(title = "Top Author Palette",
         x = "", y = "") +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          panel.grid = element_blank())

```

## Top 5 authors

For further analysis, we selected the top 5 authors in our dataset
visualizing their six most frequent colors found in their book covers.

```{r}
top_authors <- best_books  %>%
    count(author, sort = TRUE) %>% 
    arrange(desc(n))  %>% 
    slice(1)  %>%  
    pull(author)  %>%   
    head(5)  # Keep only the first 5 elements


top_authors_colors <- best_books  %>% 
    filter(author %in% top_authors) %>% 
    unnest_longer(colors) %>%  
    count(author, category, colors, sort = TRUE) %>% 
    group_by(author, category) %>%   
    slice_head(n = 6)  %>%  
    mutate(rank = row_number()) %>%  
    ungroup() %>% 
    mutate(author_label = paste0(author, " (", category, ")")) %>% 
    mutate(text_color = ifelse(
        grepl("^#", colors),  
        ifelse(
            (col2rgb(colors)[1,] * 0.299 +
             col2rgb(colors)[2,] * 0.587 +
             col2rgb(colors)[3,] * 0.114) > 150, 
            "black", "white"
        ), 
        "black"  
    ))


ggplot(top_authors_colors, aes(x = rank, y = reorder(author_label, -n), fill = colors)) +
    geom_tile(color = "gray", size = 0.5) +  
    geom_text(aes(label = colors, color = text_color), size = 2.5, fontface = "bold") +  
    scale_fill_identity() +  
    scale_color_identity() +   
    theme_minimal() +
    labs(title = "Top 6 Colors for the 5 Most Prolific Authors",
         x = "", y = "") + 
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.grid = element_blank())


```

Here it can be observed that the color palettes used by different
authors show significant variation without a clear pattern.

Regarding romance authors, Julia Quinn's covers predominantly feature
soft, pastel shades, while in contrast, Megan Maxwell's books lean
towards dark and muted tones.

Similarly, science fiction author Brandon Sanderson and Philip K. Dick
show different patterns, the former's covers mix dark and neutral shades
while the latter's books stand out with a primarily white and
light-colored palette.

This visualization provides insights into the color trends associated
with an author’s book covers, which can be useful for design analysis
and trend exploration.

# Shiny app

<https://mariacarda.shinyapps.io/data_harvesting_books/>

To examine all the analysis and the different palettes obtained, we
developed a shiny app that allows users to explore the most dominant
colors used in book covers.

```{r}

ui <- fluidPage(
  theme = shinytheme("flatly"),
  useShinyjs(),  
  
  titlePanel("The Bestseller Palette"),
  
  sidebarLayout(
    sidebarPanel(
      pickerInput("category_select", "Select a category:",
                  choices = c("All", unique(best_books$category)),
                  selected = "All", multiple = FALSE, 
                  options = list(`style` = "btn-primary")),
      
      pickerInput("author_select", "Select an author:",
                  choices = c("All"),
                  selected = "All", multiple = FALSE, 
                  options = list(`style` = "btn-info"))
    ),
    
    mainPanel(
      tabsetPanel(id = "selected_tab",
                  
        tabPanel("Book covers", fluidRow(uiOutput("book_images"))),
        
        tabPanel("Platform Palette", plotOutput("colorTiles")),
        
        tabPanel("Author Analysis", 
                 plotOutput("colorPlot"),
                 uiOutput("bookImages"))
      )
    )
  )
)

# Server
server <- function(input, output, session) {
  
  # Filtrar datos
  filtered_data <- reactive({
    data <- best_books
    
    if (input$category_select != "All") {
      data <- data %>% filter(category == input$category_select)
    }
    
    if (input$author_select != "All") {
      data <- data %>% filter(author == input$author_select)
    }
    
    return(data)
  })
  
  # Actualizar autores según categoría seleccionada
  observeEvent(input$category_select, {
    category_data <- filtered_data()
    
    updatePickerInput(session, "author_select", 
                      choices = c("All", unique(category_data$author)),
                      selected = "All")
  })
  
  # Renderizar imágenes de libros
  output$book_images <- renderUI({
    data <- filtered_data()
    
    if (nrow(data) == 0) {
      return(tags$p("No se encontraron libros que coincidan con los filtros seleccionados."))
    }
    
    image_elements <- lapply(1:nrow(data), function(i) {
      colors_list <- data$colors[[i]]
      if (is.list(colors_list)) {
        colors_list <- unlist(colors_list)
      }
      
      color_tiles <- tags$div(
        style = "display: flex; justify-content: center; margin-top: 5px;",
        lapply(colors_list, function(color) {
          tags$div(
            style = paste("background-color:", color, "; width: 30px; height: 30px; margin: 3px; border-radius: 3px;")
          )
        })
      )
      
      color_bar <- tags$div(
        style = paste("background: linear-gradient(to right, ", paste(colors_list, collapse = ", "), "); 
                      height: 30px; width: 200px; margin-top: 5px; border-radius: 3px;")
      )
      
      img_element <- tags$div(
        style = "text-align: center; margin: 5px;",
        tags$img(src = data$img_url[i], width = "150px", height = "225px")
      )
      
      combined_element <- tags$div(
        style = "width: 18%; max-width: 180px; display: flex; flex-direction: column; align-items: center; margin: 30px;",
        img_element,
        color_tiles,
        color_bar
      )
      
      combined_element
    })
    
    tags$div(
      style = "display: flex; flex-wrap: wrap; justify-content: center; 
               max-width: 100%; margin-left: auto; margin-right: auto; gap: 30px;",
      do.call(tagList, image_elements)
    )
  })
  
  # Platform Palette
  output$colorTiles <- renderPlot({
    color_data <- top_colors_all
    if (input$category_select != "All") {
      color_data <- color_data %>% filter(category == input$category_select)
    }
    
    ggplot(color_data, aes(x = factor(position), y = webpage, fill = representative_color)) +
      geom_tile(width = 0.95, height = 0.95) +
      scale_fill_identity() +
      theme_minimal() +
      labs(title = "",
           x = "", y = "") +
      theme(axis.text.x = element_blank(), 
            axis.ticks.x = element_blank(), 
            panel.grid = element_blank())
  })
  
  # Control de habilitación del selector de autor
  observeEvent(input$selected_tab, {
    if (input$selected_tab == "Platform Palette") {
      updatePickerInput(session, "author_select", selected = "All")
      disable("author_select")  # Deshabilitar el picker de autores
    } else if (input$selected_tab == "Author Analysis") {
      enable("author_select")  # Habilitar el picker de autores
    }
  })
  
  # Filtrar libros por autor
  filtered_books <- reactive({
    best_books %>% filter(author == input$author_select)
  })
  
  # Gráfico de colores
  output$colorPlot <- renderPlot({
    data <- filtered_books() %>%
      unnest_longer(colors) %>%
      count(author, colors, sort = TRUE) %>%
      slice_head(n = 6) %>%
      mutate(rank = row_number()) %>%
      ungroup() %>%
      mutate(text_color = ifelse(
        (col2rgb(colors)[1,] * 0.299 +
         col2rgb(colors)[2,] * 0.587 +
         col2rgb(colors)[3,] * 0.114) > 150, 
        "black", "white"
      ))

    ggplot(data, aes(x = rank, y = reorder(author, -n), fill = colors)) +
      geom_tile(color = "white", size = 0.5) +
      geom_text(aes(label = colors, color = text_color), size = 5, fontface = "bold") +
      scale_fill_identity() +
      scale_color_identity() +
      theme_minimal() +
      labs(title = "",
           x = "", y = "") +
      theme(axis.text.x = element_blank(), 
            axis.ticks.x = element_blank(),
            panel.grid = element_blank())
  })
  
  # Mostrar imágenes de libros
  output$bookImages <- renderUI({
    books <- filtered_books()
    img_tags <- lapply(books$img_url, function(url) {
      tags$img(src = url, width = "100px", style = "margin: 5px;")
    })
    do.call(tagList, img_tags)
  })
}

# Run the app
shinyApp(ui = ui, server = server)


```

![Screenshoot of the developed shiny
app](shiny_app.png){fig-align="center"}

# Conclusions

The analysis of book cover colors across authors, genres, and platforms
reveals no strict pattern but rather a dynamic interplay of artistic
choices. While certain genres tend to favor specific color schemes —such
as dark and cool tones for Young Adult Fantasy— there is considerable
variation between platforms. Similarly, individual authors exhibit
unique palettes, but there is no single defining scheme that dominates
across their works.

These findings suggest that book cover design is influenced by multiple
factors beyond just genre or author branding, including publisher
strategies, regional aesthetics, and reader preferences. The divergence
in colors across platforms further implies that book marketing is
tailored to different audiences, with some covers emphasizing vibrancy
while others adopt more sophisticated tones. Ultimately, it can be
concluded that visual identity is shaped by both artistic vision and
market-driven considerations.
